---
title: "Appendices"
author: "Abigail Rieck"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r packages, include = FALSE}
# Load any packages you may need here
library(car)
library(emmeans)
library(tidyverse)
library(stringr)
```


```{r readData, include = FALSE}
# Save this template file in your final project's RStudio Project 
# Edit this code chunk to read in your data

games <- read.csv("games.csv")
```

***

#### Figure 1.1: Histogram of Maximum Playtime

```{r figure1_1, echo = FALSE}
hist(games$maxplaytime,
     xlab = "Maximum Playtime (minutes)",
     main = "Histogram of Maximum Playtime")
```

***

#### Figure 1.2: Histogram of Maximum Playtime with Log Transformation

```{r figure1_2, echo=FALSE}
hist(log(games$maxplaytime),
     xlab = "log(Max Playtime in minutes)",
     main = "Histogram of log(Max Playtime)")
```

***

#### Figure 2: Histogram of Publishing Year

```{r figure2, echo=FALSE}
hist(games$yearpublished,
     xlab = "Year Published",
     main = "Histogram of Year Published")
```

***

#### Figure 3: Boxplots of Average Rating by Category

```{r figure3, echo = FALSE}
ggplot(games) +
  geom_boxplot(mapping = aes(category, average)) +
  labs(title = "Boxplots of Game Ratings by Category", x = "Game Category", y = "Average Rating")
```

***

#### Figure 4: Interaction Between Category & Age

```{r figure 4, echo = FALSE}
ggplot(games) +
  geom_jitter(mapping = aes(x = age, y = average, color = category)) +
  labs(x = "Game's Age (years)", y = "Average Rating", title = "Scatterplot of Age vs Rating by Category", color = "Category")
```

***

#### Table 1: Summary of Model 1
```{r table1, echo = FALSE}
model1 <- lm(average ~ log(maxplaytime) + log(owned)+  age*category, data = games)
summary(model1)
```

***

#### Figure 5: Residuals vs. Fitted Values Plot For Model 1

```{r figure5, echo = FALSE}
plot(model1$fitted.values, model1$residuals,
     main = "Residuals versus Fitted Values Plot \n for Model 1",
     xlab = "Fitted Values",
     ylab = "Residuals")

abline(h = 0, col = "red", lty = 2)
```

***

#### Figure 6.1: Residuals vs Fitted Values Plot for Model 2
```{r figure6_1, echo = FALSE}
model2 <- lm(average ~ log(maxplaytime) + log(owned) + age + I(age^2) + category, data = games)

plot(model2$fitted.values, model2$residuals,
     main = "Residuals versus Fitted Values Plot \n For Model 2",
     xlab = "Fitted Values",
     ylab = "Residuals")

abline(h = 0, col = "red", lty = 2)
```

***

#### Figure 6.2: QQ Plot for Model 2
```{r figure6_2, echo = FALSE}

qqnorm(model2$residuals, main = "QQ Plot of the Residuals")
qqline(model2$residuals)

```

***

#### Table 2: Summary of Final Model
```{r table2, echo = FALSE}
summary(model2)
```

***

#### Table 3: VIFs for Final Model
```{r table3, echo = FALSE}
vif(model2)
```



