---
title: "Project Work"
author: "Abigail Rieck"
date: "3/11/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(stringr)
```

```{r read in}
ratings <- read_csv("ratings.csv") 
details <- read_csv("details.csv")

                          
```
### Data Cleaning

Adding a chunk for some data cleaning- I want to combine the datasets, drop duplicated & unnecessary variables, and make a name length variable.

I will also be using code posted to tidytuesday by Alyssa Goldberg (https://rpubs.com/thewiremonkey/476630) to filter out games that have more than one category, then filter to the top 10 categories. Although I would love to include more games and categories, this is the best solution I could find for ease of analysis.

```{r data cleaning}
#join datasets
games <- left_join(details, ratings, by = "id")

#drop duplicated and unnecessary variables
games <- select(games, -`...1.x`, -`num.x`, -`...1.y`, -`num.y`, -name, -year)

#rename title variable
games <- games %>%
  rename("title" = primary)

#Create title length variable
games <- games %>%
  mutate(title_length = str_count(title))


#count number of categories games are assigned to and store as variable count
games$count <- unlist(lapply(str_split(games$boardgamecategory, ","), length))

#filter for one category only
single_cat <- games %>%
  filter(count == 1)
glimpse(unique(single_cat$boardgamecategory))
#left with 4300 games and 74 categories

#top five categories only
top_cat <- single_cat %>%
  filter(!is.na(boardgamecategory)) %>%
  group_by(boardgamecategory) %>%
  tally() %>%
  top_n(5, n) %>%
  arrange(desc(n))
sum(top_cat$n)
#still leaves us with over 2000 games to work with- plenty for the sake of this project

games <- games %>%
  filter(boardgamecategory %in% top_cat$boardgamecategory)

#make category variable more readable
games <- games %>%
  mutate(category = str_replace_all(string = boardgamecategory, pattern = "\\[\\'(.+)\\'\\]", replacement = "\\1"))

#reorder variables to make life easier
games <- games %>%
  select(title, title_length, category, yearpublished, minplaytime, maxplaytime, minplayers, maxplayers, minage, average, everything())

#Note to self- I overrode the original games data to only have games in the top 5 categories. I may regret this later if I decide to compare games outside of those categories but oh well, at least it keeps my df name short. If I need to go back and use unfiltered stuff, everything before line 37 should be rerun.

#write_csv(games, "games.csv")

```
## Project Work

### Exploratory Analysis

```{r read clean}
games <- read.csv("games.csv")
library(tidyverse)
library(stringr)
library(car)
```

```{r numericalSummaries}
#Four variables from project checkpoint:
summary(games$average)
summary(games$maxplaytime)
summary(games$minage)
summary(games$title_length)



```

```{r histograms}
##Project checkpoint variables:

#Average rating
hist(games$average,
     xlab = "Average Rating",
     main = "Histogram of Average Rating")

#Max playtime
hist(games$maxplaytime,
     xlab = "Maximum Playtime (minutes)",
     main = "Histogram of Maximum Playtime")
#Log transformation appropriate:
hist(log(games$maxplaytime),
     xlab = "log(Max Playtime in minutes)",
     main = "Histogram of log(Max Playtime)")

#Minimum age
hist(games$minage,
     xlab = "Age (years)",
     main = "Histogram of Minimum Age to Play")

#Title length
hist(games$title_length,
     xlab = "Length of Title (characters)",
     main = "Histogram of Title Length")
#Log transformation appropriate:
hist(log(games$title_length),
     xlab = "log(Title Length in characters)",
     main = "Histogram of log(Title Length)")



```

```{r scatterplotMatrix}
#For initial model
plot(games[,c(10, 6, 9, 2)])

```

```{r sideBySideBoxplot}
#For category
ggplot(games) +
  geom_boxplot(mapping = aes(category, average)) +
  labs(title = "Boxplots of Game Ratings by Category", x = "Game Category", y = "Average Rating")
  

```

```{r intitialModel}
lm_init <- lm(average ~ maxplaytime + minage + title_length + category, data = games)

summary(lm_init)
```
## Model Improvement

The model I originally build is definitely not doing much. For starters, I forgot to include the log transformations I identified, so I will do that. I also want to explore some other variables; min age doesn't have a lot of variability so it may be less interesting to look at. Year published could be fun though!

```{r new variables 1}
#First, scatterplots of transformed variables
plot(average ~ log(maxplaytime), data = games)
plot(average ~ log(title_length), data = games)

#exploring year variable
summary(games$yearpublished)
hist(games$yearpublished)

#very heavy left skew-- going to fix by changing to age!
games <- games %>%
  mutate(age = 2022 - yearpublished)
hist(games$age)
#much better! now with log transformation:
hist(log(games$age))

#write.csv(games, "games.csv")
```
To deal with the heavy left skew of the yearpublished variable, I've changed it to an age (in years, as of 2022) variable. I also overwrote my games.csv file again to include this.

Now to make some more models with this and with the log transformed original variables

First- I realized there are some problems with the log function because some of the data has 0s

#### Exploring 0s in data

```{r zeros}
summary(games$maxplaytime)
noplaytime <- filter(games, maxplaytime == 0)

#filter out 0s
games <- games %>%
  filter(maxplaytime != 0)

hist(games$maxplaytime)
hist(log(games$maxplaytime))
```
There are 74 games with a max playtime of 0. I feel like I have to assume this is some kind of error in the way the data transferred over from the site, so i will be filtering these out.
There's also some 0s in min age which probably should be considered NAs but I don't think I'll be including it in the model anyway so it should be fine

```{r log and age models}
#og variables with log transformations
lm_init_log <- lm(average ~ log(maxplaytime) + minage + title_length + category, data = games)
summary(lm_init_log)

#min age to play and log trans for title length removed
lm_noage <- lm(average ~ log(maxplaytime) + title_length + category, data = games)
summary(lm_noage)

#category removed
lm_nocat <- lm(average ~ log(maxplaytime) + title_length + minage, data = games)
summary(lm_nocat)

#age of game added to og model, no title length
lm_gameage1 <- lm(average ~ log(maxplaytime) + age + category, data = games)
summary(lm_gameage1)

#regular title length
lm_gameage2 <- lm(average ~ log(maxplaytime) + age + title_length + category, data = games)
summary(lm_gameage2)

#log title length
lm_gameage3 <- lm(average ~ log(maxplaytime) + age + log(title_length) + category, data = games)
summary(lm_gameage3)
```
Both log transformations actually made the model slightly worse. Only transforming the playtime variable made it slightly better, so I dropped the log on title length.

Based on the much lower adjusted R-squared values without playing age and category, it seems that those are both helpful predictors

Game age seemed to help the model a lot, but in this version title length is no longer a significant predictor. Gonna make some scatterplots to see whats going on

Game age 1 (no title): .2019
2 (regular title): .2023, title p-value .146
3 (log title):.2017, title p-value .2017

Title without log transformation seemed to help the model a bit but had a low p value, doesnt seem worth it -- going with game age 1


## Scatterplots part 2 electric boogaloo

```{r scatterplots two}
#Game age by category
ggplot(games) +
  geom_jitter(mapping = aes(x = age, y = average, color = category))

#title length
ggplot(games) +
  geom_jitter(mapping = aes(x = log(title_length), y = average, color = category))

#max playtime
ggplot(games) +
  geom_jitter(mapping = aes(x = log(maxplaytime), y = average, color = category))

#plot matrix
pairs( ~ average + log(maxplaytime) + log(owned) + age, data = games)

```
There does actually seem to be potential for an interaction between a game's age and category- makes sense that some may not age as well as others

```{r interaction model}

lm_interact <- lm(average ~ log(maxplaytime) + age*category, data = games)
summary(lm_interact)


```
adj R squared jumped to .2146 compared to .2049 from the model without the interaction! also fairly significant differences between the categories. definitely worth keeping, but might want to cut out other variables to make it better

### More new variables

Noticed a variable for how many people own the game that could be interesting
```{r new variables 2}
summary(games$owned)
hist(games$owned)
#massive right skew lol
hist(log(games$owned))
#much better

ggplot(games)+
  geom_point(mapping = aes(x = log(owned), y = average))

```
Looks like a really strong relationship! which makes sense; good games should get more popular

```{r owned and quadfit models}
lm_owned <- lm(average ~ log(maxplaytime) + log(owned)+  age*category, data = games)
summary(lm_owned)

lm_quadfit <- lm(average ~ log(maxplaytime) + log(owned) + age + I(age^2) + category, data = games)
summary(lm_quadfit)


```
Looking good! Adjusted R-squared is up to .2613. Still not great of course, but unless the diagnostics are bad, seems like a winner

Speaking of...

## Diagnostics

```{r residuals plot for owned}
plot(lm_owned$fitted.values, lm_owned$residuals,
     main = "Residuals versus Fitted Values Plot \n for Games with a Linear Fit",
     xlab = "Fitted Values",
     ylab = "Residuals")

abline(h = 0, col = "red", lty = 2)
```

Doesn't look great. going to scatterplot matrix a couple chunks above, it looks like there might actually be a bit of a quadratic relationship going on with log(owned), which seems way too complicated to try and fit. Let's try the model before, lm_interact

```{r residuals for interact}
plot(lm_interact$fitted.values, lm_interact$residuals,
     main = "Residuals versus Fitted Values Plot \n for Games with a Linear Fit",
     xlab = "Fitted Values",
     ylab = "Residuals")

abline(h = 0, col = "red", lty = 2)

```
Looks a little bit better but not great. Looking at the matrix again, it might make more sense to stick with log for owned and try a quadratic term for age. Only problem is I don't think I can still use my interaction term. Gonna make the model in the same chunk as owned and do residuals below.

```{r residuals quadfit}
plot(lm_quadfit$fitted.values, lm_quadfit$residuals,
     main = "Residuals versus Fitted Values Plot",
     xlab = "Fitted Values",
     ylab = "Residuals")

abline(h = 0, col = "red", lty = 2)


```
Much much better, AND the r-squared value went up! We have a new winner! Now QQ plot

```{r qqplot quadfit}
qqnorm(lm_quadfit$residuals, main = "QQ Plot of the Residuals")
qqline(lm_quadfit$residuals)

```
Beautiful! Little wonky on the tails but otherwise looks good.

### Multicollinearity & Overfitting
```{r vif}
vif(lm_quadfit)
```

No evidence of multicollinearity (low VIFs) or overfitting (no NAs in output, enough observations for predictors)

## Model Comparison
Including a chunk here to have the models I'm referencing as Model 1 and Model 2 in the report- these are the same as lm_owned and lm_quadfit (aka lm_final) above
```{r models 1 and 2}

#Model with interaction term
model1 <- lm(average ~ log(maxplaytime) + log(owned)+  age*category, data = games)
summary(model1)

#Model with quadratic term
model2 <- lm(average ~ log(maxplaytime) + log(owned) + age + I(age^2) + category, data = games)
summary(model2)

```


## Final Model

This is identical to the quadfit model above, just included here for easy access.

```{r final model}
lm_final <- lm(average ~ log(maxplaytime) + log(owned) + age + I(age^2) + category, data = games)
summary(lm_final)

```


## Summary

For the project report, I'll discuss that I played around with my original model quite a bit, looked at some variables, etc. I ultimately decided to get rid of title length bc the relationship wasn't significant even though it helped the r squared a bit. Added in age of the game (discuss transformations there too) and # of people owning it.
Originally tried out interaction between age and category and it seemed really promising until I got to the diagnostics and realized something was messed up. Going back to look more closely at the plots, I realized a quadratic fit might be more appropriate for age. This both fixed the diagnostics and improved the adjusted R-squared!


### Graphs to include

-One histogram for a log transformation
-Rating vs age by category scatterplot to discuss interaction term
-Rating vs age plot without color to discuss quadratic fit
-Residuals plot with and without quadratic fit as evidence to include it
-QQ plot for final model
